---
title: "Following along class code"
author: "Santiago Del Moral"
date: "2026-09-16"
output: pdf_document
---
Code will start here
Important to write {r} to execute R code

```{r}
x <- c(1, 2, 3)
length(x)
typeof(x)
x

```

something should've printed

```{r}
i <-c(1L, 2L, 3L, NA_integer_)
typeof(i)
i

```

We can have logical operators too
&& means AND
|| means OR
TRUE means TRUE
FALSE means FALSE
NA means NA and will return NA in operators that need it
```{r}
TRUE && TRUE
TRUE && FALSE
TRUE && NA
TRUE || NA
FALSE || NA
```

There's also matrixes use matrix()
y is the number of dimentions
use "drop=FALSE" to be able to have onedimentional matrixes instead of a vector.

```{r}
h <- matrix(1:10, 1)
h
```


there are lists that can contain any data type
data frames are two dimentional data structures
a $ can help you fetch
```{r}
df <- data.frame(a=1:10, b=11:20, c=letters[1:10])
df
df$a
```

Functions

R is a functional programming language


```{r}
add <-function(x,y) {
  x+y
}

add(1,2)

```

more functions
apparently you can type " \ " instead of function
In R things are considered pass by value not pass by reference
so x is considered immutable... 
which means unless you're doing <- then only the function's copy is being updated
Original remains unchanged, unless assigned to a new variable

```{r}
x <- 1:10
foo <- function(x){
  x[1] <- -99
  x
}

foo(x)
#we'll see x is unchanged
x
x<-foo(x)
x
#now x is changed because it's been assigned
```
 
No need to keep track of mutations and changes because of immutability.

BREAK

Data visualization now
Why do we look at data? 
because the summary statistics can be misleading sometimes
can check intuitively for outliers
find patterns visually

Pie charts are not the best usually, plain bar charts are better
QQ plot, a empirical in the y axis and the theoretical in the x axis (usually normal dist)

Looking at multiple variables, how to check relationships?
scatterplot
jitter to avoid overplotting in a scatter plot can work to break up things

density plots 1D or 2D are cool

Plotting:
we'll be using library(ggplot2) for now


```{r}
library(ggplot2)
ggplot(penguins,
       aes(x=bill_len,y=body_mass))+geom_point()

```

```{python}
print("helloworld")
```

